Microplastics in the Urban Environment

### 2021-10-26 products/manuscript/Manuscript.Rmd
fixed manuscript citation style

---
title: "Microplastics in the Urban Environment"
subtitle: "Exploratory Study of Freshwater Microplastics in Athens, GA"
author: Amelia M. Foley
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: article
output:
  bookdown::word_document2: 
    toc: false
    number_sections: true
  bookdown::html_document2: 
    toc: false
#bibliography: library.bib #I am having trouble with the bib file (error message: pandoc error 83 file not found). Removing this line for now so that Manuscript will knit. Will continue to troubleshoot and fix bib formatting. 
csl: ../apa.csl
---

@@

Through decades of research, the scientific community proves that microplastics are ubiquitous, having identified microplastics in oceans, rivers, lakes, and other water bodies across the globe, including remote mountain lakes, the bottom of the ocean, and in ocean sediments offshore of Antarctica [@Free2014][@Waller2017][@Peng2018]. Modeling of current and future plastic production and pollution indicates that plastics are so pervasive that they have been considered as a geological indicator of the proposed Anthropocene era [@Zalasiewicz2016]. These models also suggest that microplastic concentrations will only increase in the future [@Geyer2017][@Barboza 2018][@Prokic2019][@Evereart2018]. In a 2018 paper, Barboza and collaborators state that the increase in environmental microplastics is particularly concerning due to the contaminants’ small size, the limited technology for quantifying their presence, and their potential for adversely affecting both marine biota and humans [@Barboza2018]. Further, research shows that microplastic pollution is largely due to land-based sources; ocean-based sources such as commercial fishing, vessels, and other activities, contribute only 20% of the total plastic debris, while terrestrial activity contributes the other 80% of pollution in the marine environment [@Andrady2011][@Li2018]. Microplastics with various terrestrial origins mainly enter the marine environment via rivers, releasing most microplastics to the ocean and retaining some in freshwater systems [@Free2014][@Li2018][@Browne2010][@Lebreton2017]. However, a majority of microplastics research to date has focused on marine settings rather than freshwater, indicating that more studies are required in this area [@Besseling2018][@Dris2015]. Examining freshwater systems as the dominant source of microplastics will lead to a better overall understanding of microplastics pollution input and therefore insight into the scale of the issue and mitigation strategies [@Lebreton2017].

>>>>

---
title: "Microplastics in the Urban Environment"
subtitle: "Exploratory Study of Freshwater Microplastics in Athens, GA"
author: Amelia M. Foley
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: article
output:
  bookdown::word_document2: 
    toc: false
    number_sections: true
  bookdown::html_document2: 
    toc: false
bibliography: library.bib #I am having trouble with the bib file (error message: pandoc error 83 file not found). Removing this line for now so that Manuscript will knit. Will continue to troubleshoot and fix bib formatting. Bib file works now, but some citations recieve double brackets. Will continue looking into this issue
csl: elsevier.csl
---

@@

Through decades of research, the scientific community proves that microplastics are ubiquitous, having identified microplastics in oceans, rivers, lakes, and other water bodies across the globe, including remote mountain lakes, the bottom of the ocean, and in ocean sediments offshore of Antarctica [@Free2014;@Waller2017;@Peng2018]. Modeling of current and future plastic production and pollution indicates that plastics are so pervasive that they have been considered as a geological indicator of the proposed Anthropocene era [@Zalasiewicz2016]. These models also suggest that microplastic concentrations will only increase in the future [@Geyer2017;@Barboza2018;@Prokic2019;@Everaert2018]. In a 2018 paper, Barboza and collaborators state that the increase in environmental microplastics is particularly concerning due to the contaminants’ small size, the limited technology for quantifying their presence, and their potential for adversely affecting both marine biota and humans [@Barboza2018]. Further, research shows that microplastic pollution is largely due to land-based sources; ocean-based sources such as commercial fishing, vessels, and other activities, contribute only 20% of the total plastic debris, while terrestrial activity contributes the other 80% of pollution in the marine environment [@Andrady2011;@Li2018]. Microplastics with various terrestrial origins mainly enter the marine environment via rivers, releasing most microplastics to the ocean and retaining some in freshwater systems [@Free2014;@Li2018;@Browne2010;@Lebreton2017]. However, a majority of microplastics research to date has focused on marine settings rather than freshwater, indicating that more studies are required in this area [@Besseling2017;@Dris2015]. Examining freshwater systems as the dominant source of microplastics will lead to a better overall understanding of microplastics pollution input and therefore insight into the scale of the issue and mitigation strategies [@Lebreton2017].
